---
title: "EDA"
subtitle: "Part III"
output: github_document
---

```{r setup}
library(tidyverse)
library(patchwork)
```

## Load the weather data


```{r}
library(p8105.datasets)
data("weather_df")
  
weather_df = weather_df %>%
  mutate(month = floor_date(date, unit = "month"))
```

## `group by`

```{r}
weather_df %>% 
  group_by(name, month) 
```

## counting things

count month observations

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(n_obs = n())
```

we can use `count()`

```{r}
weather_df %>% 
  count(name, month, name = "n_obs")
```

**NEVER** use base R's `table`

```{r, eval = FALSE}
weather_df %>% 
  pull(month) %>% 
  table()
```

other helpful counters

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    n_obs = n(),
    n_days = n_distinct(date)) 
```

## A digression on 2x2 tables

```{r}
weather_df %>% 
  filter(name != "Molokai_HI" ) %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not cold",
      TRUE      ~ "" 
    )
  ) %>% 
  group_by(name, cold) %>% 
  summarize(count = n())
```

```{r}
weather_df %>% 
  filter(name != "Molokai_HI" ) %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not cold",
      TRUE      ~ "" 
    )
  ) %>% 
  janitor::tabyl(name, cold)
```


## General Summaries

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  )
```

This is a dataframe!

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()
```

Suppose you want to summarize many columns
`across` is to list multiple things across same function

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(across(prcp:tmin, mean))
```

Reminder: sometimes results are easier to read in another format

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(
    names_from = name,
    values_from = mean_tmax
  ) %>% 
  knitr::kable(digits = 1)
```


